Article Title

### 2013-11-30 plos.tex
Updated Abstract, Authors' Summary, and Introduction.

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Correlations between the spiking activity of pairs of neurons are among the most familiar descriptive statistics of neural activity.
Multineuronal recordings provide richer information than the equivalent number of neuron pairs.
For example, the full covariance  matrix reveals correlated activity across the entire population as well as partial correlations between pairs.
Estimation of covariance matrices can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
Optimal regularization must be determined empirically since its effect depends on how closely the imposed structure matches the underlying regularities.
For example, we can use low-rank parameterizations of the covariance matrix to account for common fluctuations across the recorded population.
Conversely, if correlations are strongly influenced by a small fraction of pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix.

We can also use a `sparse+low' rank inverse covariance representation to account for both common fluctuations and pairwise interactions.  
To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance. We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.  Each covariance estimator was optimized and evaluated by cross-validation. As expected, covariance shrinkage reliably outperformed the sample covariance estimate. In turn, factor analysis-based estimates significantly outperformed covariance shrinkage. Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage estimators. The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in the circuit we studied.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
Covariance matrices of the activity of populations of neurons are useful descriptions of the functional organization of neural circuits with implications for stimulus coding and neural circuit architecture. 
Estimates of the covariance matrix can be dramatically improved by \emph{regularization}, \emph{i.e.}\;the biasing of the usual estimate toward one of several possible low-dimensional approximations, the \emph{target estimate}. However, the amount of improvement depends on how closely the selected approximation reflects the dominant interactions in the circuit. 
In simulation, we demonstrated that no regularized estimate was universally superior to others: each performed best when the target estimate was the closest match to the structure of the data generating process. 
We then evaluated covariance matrix estimates on the calcium activity recorded \emph{in vivo} from large populations of closely spaced neurons  in mouse primary visual cortex, after subtracting the average stimulus response.
Using cross-validation, we found that the best estimate was produced when the covariance matrix was approximated as that of a sparse Gaussian graphical models with a small number of latent units. 
Thanks to the paucity of their parameters, these optimally estimated covariance matrices also help interpretation of the covariance structure and allow relating the functional organization of the neural circuit to its cytoarchitecture.

\section*{Introduction}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.  
For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.  
Noise correlations can have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}, and have been interpreted to indicate detailed and specific functional organization. 
Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity and stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, cell-type specificity (?), progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

However, neural correlations do not come with ready or unambiguous mechanistic interpretations.
Theoretical and simulation studies have shown that neural correlations, rise at various temporal scales from combinations of multiple underlying mechanisms.
These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.

Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \cite{Shadlen:1998,Zohary:1994}.
Multineuronal recordings allow estimation of covariance matrices of large populations of neurons.
Such estimates provide more information than the equivalent number of pairwise correlations assessed in isolation.
Indeed, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations that accentuate different aspects of the correlation structure and may suggest different mechanistic interpretations.
For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for the activity of all the other recorded cells; a strong interaction between a pair of neurons may be expressed by a single partial correlation but its effects may propagate to multiple correlations and eigenvalues.
The inverse of the covariance matrix plays an important role in decoding schemes such as linear discriminant analysis, for example.
The mutliple  representations of the covariance matrix with their alternative interpretations add both complications and opportunities into the search for fundamental regularities in neural population activity.

With large numbers of recorded cells, the usual estimations of the covariance matrix from finite recordings become ill-conditioned. Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically. 
For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Ledoit:2004}. Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.

With this study, we pursue two related goals: 
\begin{enumerate}
	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
	\item Aid interpretation of neural covariance matrices.
\end{enumerate}

To select the best covariance matrix estimate for a specific neural circuit, we evaluate the performance of four estimators using different regularization schemes, each motivated by  a different hypothesis for the origin of neural correlations. \emph{Regularization} is the deliberate biasing the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \cite{Schafer:2005,Bickel:2006}.
Regularized estimators allow reaching a favorable tradeoff between bias and variance in order to minimize the total error.

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.  It is anticipated that correlations measured in such recordings will uncover aspects of functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings can be challenging.  Estimation can be improved by regularization: the biasing of the estimate toward a low-dimensional, less variable approximation.  The amount of improvement depends on how closely the reduced approximation captures the dominant interactions with fewest terms.  Therefore, the selection of the most efficient estimator is an empirical question dependent on the system under investigation.

We compared regularized correlation matrix estimators biased toward four respective families of low-dimensional correlation structures: independent, latent factors, sparse partial correlations, and sparse partial correlations with latent factors.  The estimators' performance was evaluated on the noise correlation matrices from spatially compact groups of 91--314 neurons (average 206) in mouse visual cortex (31 sites in 24 animals) acquired by \emph{in vivo} fast 3D random-access two-photon imaging of calcium signals.  Recordings lasting between 15 and 20 minutes were deconvolved and binned at 150 ms.  Each estimator was optimized and evaluated by nested cross-validation.  We found that both sparse partial correlations and latent factors were required for efficient estimation.  In addition to reducing estimation error, optimized estimates provided compact representations of the correlation structure consisting of a sparse network of pairwise interactions and global diffuse fluctuations.  The density of positive interactions in these networks decreased rapidly with distance and with difference in orientation preference, whereas negative interactions were less selective.  The physiological interpretation of these components remains in question and must be addressed by future experiments where synaptic connectivity, cell types, and local field potentials may be used to corroborate such interpretations.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlation matrices of the activity of populations of neurons are useful descriptors of their functional organization with implications for stimulus coding and circuit architecture.  Estimates of the correlation matrix can be dramatically improved by \emph{regularization}: the biasing of the estimate toward one of several possible low-dimensional approximations.  Using synthetic data, we demonstrated that different estimators performed better or worse than others depending on how well their regularization targets captured the structure of the true correlation matrix.  Then, we compared the performance of several estimators on the population calcium activity of relatively large and dense groups of neurons in mouse visual cortex under visual stimulation.  We found that the best estimator for such neural data was composed of a sparse network of partial correlations between pairs of neurons combined with several latent units interacting with the entire population. We hypothesize that such succinct representations of the correlation structure may more accurately reflect the circuit's functional organization than the conventional correlations.  
\section*{Introduction}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect aspects of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.  Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.

A correlation matrix provides more information than the equivalent number of pairwise correlations assessed in isolation.  Yet early studies of neural correlations were based on measurements from isolated pairs of neurons; the effects of correlations on neural coding were then extrapolated in simulations and theoretical studies \cite{Shadlen:1998,Zohary:1994}.  Modern multineuronal recordings from increasingly large and dense subsets of neurons in intact circuits allow estimating the entire correlation matrix of large populations of highly interconnected neurons.  Such estimates can be transformed into various representations that express different aspects of the correlation structure and may suggest different mechanistic interpretations.  For example, factor analysis or principle component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.  The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.  Under several assumptions (predominance of linear effects, sufficiently complete sampling of interacting neurons, and sufficient amount of data), partial correlations may better reflect the direct statistical association between pairs of neurons.

Large sizes of recorded populations present major challenges to the estimation and interpretation of correlations.  As the population size increases, the amount of recorded data grows only linearly whereas the number of estimated coefficients grows quadratically.  This mismatch leads to increased opportunities for spurious correlations, overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.

In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.

Both aims benefit from \emph{regularization}: the deliberate biasing (\emph{`shrinkage'}) of the estimate toward one of several possible low-dimensional approximations called here \emph{target estimates} \cite{Schafer:2005,Bickel:2006}. Low-dimensional estimates are less susceptible to sampling noise in the empirical data and are thus less variable.  However, they are also biased away from the true correlation matrix toward its projection onto the space of low-dimensional approximations. In what is called \emph{Stein's paradox}, arbitrary low-dimensional targets can be used to reduce
